---
output: github_document
---

# ContextLinkr

ContextLinkr is an R package for linking individual-level records to geographic
contextual data for multilevel health research.

The package is currently in early development. The first implemented function is
`gc_address()`, which geocodes address data using configurable geocoding
services.

## Installation

```{r eval=FALSE}
# Not yet available from CRAN.
# Development installation instructions will be added after the GitHub
# repository is public.
```

## Basic geocoding example

`gc_address()` accepts either a full address column or separate street, city,
state, and ZIP code columns.

Because geocoding may send address information to external services, users must
explicitly confirm that they have permission to do so.

```{r eval=FALSE}
library(ContextLinkr)

res <- gc_address(
  sample_addresses,
  street = street,
  city = city,
  state = state,
  zip = zip,
  geocoder = "osm",
  confirm_external = TRUE
)

res
geocode_summary(res)
geocode_successes(res)
geocode_failures(res)
```

Users working with protected, sensitive, or restricted data should geocode only
in approved computing environments. If coordinates are already available,
downstream ContextLinkr functions will allow users to skip geocoding and proceed
directly to geographic and contextual linkage steps.

## Census tract identification example

`id_tract()` identifies the Census tract containing each coordinate pair.

```{r eval=FALSE}
dc_test <- tibble::tibble(
  id = 1,
  latitude = 38.8977,
  longitude = -77.0365
)

dc_tract <- id_tract(
  dc_test,
  lat = latitude,
  lon = longitude,
  state = "DC"
)

dc_tract
tract_summary(dc_tract)
tract_successes(dc_tract)
tract_failures(dc_tract)
```

`state` can contain one or more states when coordinates span multiple states.

Rows with missing latitude or longitude are retained in the output but receive
missing tract fields and `.tract_identified = FALSE`.

```{r eval=FALSE}
regional_test <- tibble::tibble(
  id = 1:2,
  latitude = c(38.8977, 39.2904),
  longitude = c(-77.0365, -76.6122)
)

regional_tracts <- id_tract(
  regional_test,
  lat = latitude,
  lon = longitude,
  state = c("DC", "MD")
)

regional_tracts
tract_summary(regional_tracts)
```

By default, `id_tract()` enables `tigris` caching during the call so repeated
tract lookups do not require repeated boundary downloads.

## Link records to Census tracts

`link_context()` provides an end-to-end wrapper for linking individual-level
records to Census tract geography. If latitude and longitude columns are already
available, `link_context()` skips geocoding and identifies Census tracts directly.

```{r link-context-coordinates, eval=FALSE}
linked <- link_context(
  sample_addresses,
  address = address,
  state = "DC",
  geocoder = "census_single",
  confirm_external = TRUE
)

linked
```

You can summarize the linked output with `link_summary()`:

```{r link-summary, eval=FALSE}
link_summary(linked)
```

You can also separate successfully linked records from records that still need
review:

```{r link-filters, eval=FALSE}
linked_records <- link_successes(linked)
records_to_review <- link_failures(linked)
```

These helpers are useful for basic QA after geocoding and tract identification.

## Join contextual variables

After records have been linked to Census tract geography, `join_context()` can
join tract-level contextual variables to individual-level records.

```{r join-context, eval=FALSE}
context <- tibble::tibble(
  tract_geoid = c("11001980000", "24510040100"),
  deprivation_index = c(0.8, 0.6),
  rurality = c("urban", "urban")
)

joined <- join_context(
  linked,
  context,
  by = "tract_geoid"
)

joined
```

The output includes `.context_joined`, which indicates whether each linked
record matched a row in the contextual dataset.

`join_context()` does not download contextual data. It joins contextual variables
that are already available in memory.

## Development roadmap

Planned core functions include:

- `gc_address()` — geocode address data.
- `id_tract()` — identify Census tracts containing coordinates.
- `get_context()` — retrieve contextual variables from Cancer InFocus.
- `link_context()` — join individual records to contextual variables through a
  wrapper workflow.
