---
title: "ContextLinkr workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ContextLinkr workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

This vignette demonstrates the main ContextLinkr workflow for linking
individual-level records to place-based contextual measures. In the typical
workflow, users start with individual-level records, identify Census tract GEOIDs
through address geocoding and tract linkage when needed, and then add hosted
Cancer InFocus contextual measures with `add_context()`.

ContextLinkr also includes lower-level helpers for joining user-supplied
contextual datasets, but those helpers are intended for advanced or transitional
workflows. Most users should rely on Cancer InFocus context retrieval so
contextual measures, source metadata, cache behavior, and output provenance are
handled consistently.

The typical workflow is:

1. Start with individual-level records that contain addresses, latitude/longitude
   coordinates, or existing Census tract GEOIDs.
2. Identify Census tract GEOIDs with `link_context()` or `id_tract()` when tract
   GEOIDs are not already available.
3. Review linkage results with summary, success, and failure helpers.
4. Add selected Cancer InFocus contextual measures with `add_context()` or
   `link_context(include_context = TRUE)`.
5. Inspect source metadata, cache information, and output provenance.

This vignette uses `eval = FALSE` examples so that it does not send address data
to external services or retrieve hosted data during package checks.

ContextLinkr is under active development. Function names, arguments, and output
structures may change before a stable release. For collaborator testing, use a
tagged release or a specific Git commit when possible.

## Start with individual-level records

For beta testing and examples, use synthetic records or public geography
identifiers. Do not use patient, participant, or identifiable address-level data
when testing package installation or basic workflows.

```{r example-records}
library(ContextLinkr)

records <- data.frame(
  person_id = c("test_001", "test_002"),
  tract_geoid = c("21067003600", "21067004205")
)

records
```

If your records already contain Census tract GEOIDs, you can skip geocoding and
tract identification and go directly to `add_context()`.

## Link records to Census tracts

If records contain latitude and longitude, `link_context()` can identify Census
tracts without sending address fields to a geocoder.

```{r link-coordinates}
coordinate_records <- data.frame(
  person_id = "test_001",
  latitude = 38.8977,
  longitude = -77.0365
)

linked <- link_context(
  coordinate_records,
  lat = latitude,
  lon = longitude,
  state = "DC"
)

linked
```

If records contain address fields instead of coordinates, `link_context()` can
geocode records before tract identification. Because this sends address data to
an external geocoding service, `confirm_external = TRUE` is required.

```{r link-addresses}
address_records <- data.frame(
  person_id = "test_001",
  address = "1600 Pennsylvania Ave NW, Washington, DC 20500"
)

linked_from_address <- link_context(
  address_records,
  address = address,
  state = "DC",
  geocoder = "census_single",
  confirm_external = TRUE
)

linked_from_address
```

## Review linkage results

After records are linked, use the summary helpers to inspect the result.

```{r linkage-summary}
link_summary(linked)
link_successes(linked)
link_failures(linked)
```

## Find Cancer InFocus contextual measures

Use `available_context_measures()` to review available Cancer InFocus contextual
measure definitions.

```{r available-measures}
available_context_measures()
```

Use `search_context_measures()` to search measure metadata by keyword.

```{r search-measures}
search_context_measures("population")
```

The `def` column contains the user-facing measure definitions to pass to
`measures` or `context_measures`.

## Add Cancer InFocus context to records

If records already contain a tract GEOID column, use `add_context()` to retrieve
Cancer InFocus contextual measures and join them to the records.

```{r add-context}
records_with_context <- add_context(
  .data = records,
  tract_col = "tract_geoid",
  measures = c("Total Population", "Living Below Poverty"),
  use_cache = TRUE,
  refresh_cache = FALSE
)

records_with_context
```

Review context-joining results with the context summary helpers.

```{r context-summary}
context_summary(records_with_context)
context_successes(records_with_context)
context_failures(records_with_context)
```

Inspect provenance for the contextual measures that were added.

```{r context-provenance}
context_provenance(records_with_context)
```

## Run the end-to-end workflow in one call

For convenience, `link_context()` can also add Cancer InFocus context after tract
linkage.

```{r end-to-end}
linked_with_context <- link_context(
  coordinate_records,
  lat = latitude,
  lon = longitude,
  state = "DC",
  include_context = TRUE,
  context_measures = "Total Population"
)

linked_with_context

context_provenance(linked_with_context)
```

## Retrieve context directly

For advanced workflows, `get_context()` retrieves Cancer InFocus context for
specific geography IDs.

```{r get-context}
tract_context <- get_context(
  geographies = c("21067003600", "21067004205"),
  measures = "Total Population",
  geography = "tract",
  format = "wide",
  use_cache = TRUE,
  refresh_cache = FALSE
)

tract_context
```

Use `format = "long"` when you want metadata-rich rows suitable for inspection or
analysis, and `format = "wide"` when you want one row per geography for joining.

```{r get-context-long}
tract_context_long <- get_context(
  geographies = c("21067003600", "21067004205"),
  measures = "Total Population",
  geography = "tract",
  format = "long",
  use_cache = TRUE,
  refresh_cache = FALSE
)

tract_context_long
```

## Inspect cache and source metadata

ContextLinkr can cache hosted Cancer InFocus context files locally. Cache helpers
are useful for troubleshooting and for confirming what has been downloaded.

```{r cache-info}
context_cache_info()
```

Use `context_data_sources()` to inspect hosted source metadata.

```{r context-data-sources}
context_data_sources()
```

To force a fresh download of hosted context data, use `refresh_cache = TRUE`.

```{r refresh-cache}
refreshed_context <- get_context(
  geographies = "21067003600",
  measures = "Total Population",
  geography = "tract",
  format = "wide",
  use_cache = TRUE,
  refresh_cache = TRUE
)

refreshed_context
```

To clear the local ContextLinkr cache, use explicit confirmation.

```{r clear-cache}
clear_context_cache(confirm = TRUE)
```

## Advanced: join a prepared context dataset

Most users should use `add_context()` to retrieve and append hosted Cancer
InFocus contextual measures. The lower-level `join_context()` helper is available
when users already have a prepared contextual dataset with one row per geographic
key.

```{r join-context}
prepared_context <- data.frame(
  tract_geoid = c("21067003600", "21067004205"),
  local_context_score = c(0.25, 0.75)
)

joined <- join_context(
  records,
  context = prepared_context,
  by = "tract_geoid"
)

joined
```

## Privacy and data flow

ContextLinkr keeps individual-level records local unless a user explicitly
requests address-based geocoding through an external service.

Cancer InFocus context retrieval uses geography identifiers, such as Census tract
GEOIDs, to retrieve public contextual measures. Users should not send names,
medical record numbers, dates of birth, street addresses, or other direct
personal identifiers to Cancer InFocus context retrieval functions.

For beta testing, do not use real patient, participant, or identifiable data.
Use synthetic records or public example tract GEOIDs.
