---
title: "ContextLinkr workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ContextLinkr workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

ContextLinkr supports workflows that link individual-level records to place-based
contextual measures. The typical workflow is:

1. Start with records that contain addresses, latitude/longitude coordinates, or
   Census tract GEOIDs.
2. Identify Census tract geography with `link_context()` or `id_tract()`.
3. Add selected Cancer InFocus contextual measures with `add_context()` or
   `link_context(include_context = TRUE)`.
4. Review linkage and context-joining results with summary helpers.

This vignette uses `eval = FALSE` examples so that it does not send address data
to external services or retrieve hosted data during package checks.

## Link records to Census tracts

If records already contain latitude and longitude, `link_context()` can identify
Census tracts without sending address fields to a geocoder.

```{r link-coordinates}
library(ContextLinkr)

records <- tibble::tibble(
  id = 1,
  latitude = 38.8977,
  longitude = -77.0365
)

linked <- link_context(
  records,
  lat = latitude,
  lon = longitude,
  state = "DC"
)

linked
```

If records contain address fields instead of coordinates, `link_context()` can
geocode records before tract identification. Because this sends address data to
an external geocoding service, `confirm_external = TRUE` is required.

```{r link-addresses}
linked_from_address <- link_context(
  sample_addresses,
  address = address,
  state = "DC",
  geocoder = "census_single",
  confirm_external = TRUE
)

linked_from_address
```

## Review linkage results

After records are linked, use the summary helpers to inspect the result.

```{r linkage-summary}
link_summary(linked)
link_successes(linked)
link_failures(linked)
```

## Find Cancer InFocus contextual measures

Use `available_context_measures()` to review available Cancer InFocus contextual
measure definitions.

```{r available-measures}
available_context_measures("tract")
```

Use `search_context_measures()` to search measure metadata by keyword.

```{r search-measures}
search_context_measures("population", geography = "tract")
```

The `def` column contains the values to pass to `measures` or
`context_measures`.

## Add context to linked records

If records already contain a tract GEOID column, use `add_context()` to retrieve
Cancer InFocus contextual variables and join them to the records.

```{r add-context}
linked_with_context <- add_context(
  linked,
  tract_col = tract_geoid,
  measures = "Total Population"
)

linked_with_context
```

Review context-joining results with the context summary helpers.

```{r context-summary}
context_summary(linked_with_context)
context_successes(linked_with_context)
context_failures(linked_with_context)
```

## Run the end-to-end workflow in one call

For convenience, `link_context()` can also add Cancer InFocus context after tract
linkage.

```{r end-to-end}
linked_with_context <- link_context(
  records,
  lat = latitude,
  lon = longitude,
  state = "DC",
  include_context = TRUE,
  context_measures = "Total Population"
)

linked_with_context
```

## Retrieve context directly

For advanced workflows, `get_context()` retrieves Cancer InFocus context for
specific geography IDs.

```{r get-context}
tract_context <- get_context(
  geographies = "11001006202",
  measures = "Total Population",
  geography = "tract",
  format = "wide"
)

tract_context
```

Use `format = "long"` when you want metadata-rich rows suitable for inspection or
analysis, and `format = "wide"` when you want one row per geography for joining.

## Privacy and data flow

ContextLinkr keeps individual-level records local unless a user explicitly
requests address-based geocoding through an external service.

Cancer InFocus context retrieval uses geography identifiers, such as Census tract
GEOIDs, to retrieve public contextual measures. Users should not send names,
medical record numbers, dates of birth, street addresses, or other direct
personal identifiers to Cancer InFocus context retrieval functions.
