---
title: "Joining User-Supplied Contextual Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Joining User-Supplied Contextual Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

ContextLinkr links individual-level records to geographic contextual data for
multilevel health research.

This vignette demonstrates the workflow for joining linked records to a
user-supplied contextual dataset. The workflow assumes contextual variables are
already available in memory. ContextLinkr does not download contextual data in
this step.

The basic workflow is:

1. Link individual-level records to Census tract geography.
2. Join tract-level contextual variables to the linked records.
3. Summarize context join completeness.
4. Separate successful joins from records that need review.
5. Identify linked tract GEOIDs that are missing from the contextual dataset.

# Example data

ContextLinkr includes two small example datasets:

- `sample_addresses`: example individual-level address records.
- `sample_context`: example tract-level contextual variables.

```{r}
library(ContextLinkr)

sample_addresses
sample_context
```

# Link records to Census tracts

In a full workflow, `link_context()` can geocode address records and identify
Census tracts. Because geocoding may send address information to external
services, users must explicitly confirm that they have permission to geocode
those records.

This example is not evaluated in the vignette because it may require external
services and downloaded tract boundaries.

```{r eval = FALSE}
linked <- link_context(
  sample_addresses,
  address = address,
  state = "DC",
  geocoder = "census_single",
  confirm_external = TRUE
)

linked
link_summary(linked)
```

For the remaining examples, we will create a small linked dataset directly.
This keeps the vignette lightweight and avoids external service calls.

```{r}
linked <- tibble::tibble(
  id = 1:4,
  tract_geoid = c(
    "11001980000",
    "11001980100",
    "11001980200",
    "99999999999"
  )
)

linked
```

# Join contextual variables

Use `join_context()` to join tract-level contextual variables to linked records.

```{r}
joined <- join_context(
  linked,
  sample_context,
  by = "tract_geoid"
)

joined
```

The output preserves the individual-level records and adds contextual variables
from `sample_context`. It also adds `.context_joined`, which indicates whether
each linked record matched a row in the contextual dataset.

# Summarize context join completeness

Use `context_summary()` to summarize the context join.

```{r}
context_summary(joined)
```

If geocoding and tract-identification metadata are also present, `link_summary()`
can summarize the broader linkage workflow.

```{r}
link_summary(joined)
```

# Extract successful and unsuccessful context joins

Use `context_successes()` to retain records that matched contextual data.

```{r}
context_successes(joined)
```

Use `context_failures()` to retain records that did not match contextual data.

```{r}
context_failures(joined)
```

# Identify missing contextual keys

Use `missing_context_keys()` to list tract GEOIDs that appear in the linked
records but are absent from the contextual dataset.

```{r}
missing_context_keys(linked, sample_context)
```

This is useful for quality assurance before or after joining contextual data.

# Different key names

If the linked records and contextual dataset use different join-key names, pass
a named character vector to `by`. The name is the key in the linked records, and
the value is the key in the contextual dataset.

```{r}
context_alt_key <- sample_context
names(context_alt_key)[names(context_alt_key) == "tract_geoid"] <- "GEOID"

joined_alt_key <- join_context(
  linked,
  context_alt_key,
  by = c("tract_geoid" = "GEOID")
)

joined_alt_key
```

# Notes for sensitive data

Users working with protected, sensitive, or restricted data should complete
geocoding and geographic linkage only in approved computing environments.
ContextLinkr is designed so users can join contextual data after geographic
linkage has already occurred.
